---
title: "Data Manipulation"
author: "Xiaoyu Wu"
date: "2023-09-17"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in the FAS Litters Data

```{r }
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
pups_df = read_csv("./data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

## `select`

Choose some columms and not others 

```{r}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)
```
You can specify a range of columns to keep, from group column to gd_of_birth column 
```{r}
select(litters_df, group:gd_of_birth)
```
You can also specify columns you’d like to remove:
```{r}
select(litters_df, -pups_survive)
```
You can rename variables as part of this process:
```{r}
select(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
rename(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```
helper functions for select 
everything() function helps select everything else after litter_number and pups_survive
```{r}
select(litters_df, starts_with("gd"))
select(litters_df, litter_number, pups_survive, everything())
relocate(litters_df, litter_number, pups_survive)
```
skip a column named litter_number
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```
`select` vs `pull`
```{r}
select(litters_df, group)
pull(litters_df, group)
```

## `filter`

```{r}
filter(litters_df,gd0_weight<22 )
```

```{r}
filter(litters_df,gd0_weight>=22 )
```

```{r}
filter(litters_df,gd_of_birth==20 )
```

```{r}
filter(litters_df,gd_of_birth!=20 )
```

```{r}
filter(litters_df,gd0_weight>=22,gd_of_birth==20 )
```
group including either Con7 or Mod8 
```{r}
filter(litters_df,group %in% c("Con7","Mod8") )
```
short `select` learning assessment 
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `mutate`

```{r}
mutate(litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## `arrange`

```{r}
arrange(litters_df, group, pups_born_alive)
```

## pipe `%>%` cmd+shift+m 

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_df
```

