---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in some data 

Read in the litters dataset. 
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```

Read in pups dataset 
```{r}
pups_data = read_csv(file = "./data/FAS_pups.csv")
pups_data = read_csv(file = "~/Desktop/Data Science/script and data/data_wrangling_i/data/FAS_litters.csv")
pups_data = janitor::clean_names(pups_data)
pups_data
```

## Take a look at the data

Printing in the console. 

```{r}
litters_df
```
See part of the dataset 
```{r}
head(litters_df)
tail(litters_df)
```
view functions to see all the dataset
```{r eval=FALSE}
view(litters_df)
```
Look at data summary including variables types 
```{r}
str(litters_df)
```
Look at dataset using `skim` function 
```{r}
skimr::skim(litters_df)
```


## Options to read_csv

check out ` ?read_csv()` for more information. 
```{r}
litters_df=read_csv("./data/FAS_litters.csv", skip=10, col_names=FALSE)
litters_df
```

adding `NA` values 
```{r}
litters_df=read_csv("./data/FAS_litters.csv", na=c("NA",19))
litters_df
```
changing columns type treatment 
```{r}
litters_df=read_csv("./data/FAS_litters.csv", col_types = cols(Group=col_factor()))
litters_df
```

## Other file format

Read in an excel file.

```{r}
mlb_df=read_excel("./data/mlb11.xlsx", range="A1:F7")
mlb_df
```

Read in a SAS file.
```{r}
pulse_df=read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with Base R

what about `read.csv` ...? 

```{r eval=FALSE}
litters_base=read.csv("./data/FAS_litters.csv")
litters_readr=read_csv("./data/FAS_litters.csv")
litters_base
litters_readr
```

## Exoprting data

Export the mlb sub-table
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df_cleaned = janitor::clean_names(litters_df)
write_csv(litters_df_cleaned,"data/litters_cleaned.csv")
```

